---
title: "Sample analysis"
format: docx
---

```{r setup, include=FALSE}
# Load packages:

library(knitr)
library(here)
library(readr)
library(dplyr)
library(forcats)
library(stringr)
library(tidyr)
library(labelled)
# library(gtsummary) # TODO: Uncomment (& install packages if using {gtsummary})
# library(broom)
# library(broom.helpers)
library(survival)


# Environment configuration:

## Set working directory to the project root (reproducibly):
opts_knit$set(root.dir = here())

# Set chunk options:
opts_chunk$set(echo = FALSE) # Omit syntax from output
```

# Introduction

We will replicate the data analysis in the [`nwtco` help
page](https://www.rdocumentation.org/packages/survival/versions/3.5-5/topics/nwtco)
of the R [{survival} package](https://cran.r-project.org/package=survival)
v3.5-5.

# Methods

We load and preprocess the dataset, and then compute their descriptive
statistics and the contingency table of the cancer histologies. Finally, we fit
a Cox proportional hazards model to the data.

```{r preprocess-data}
## Read dataset:
study_data <- read_csv(
  # TODO: Change "<dataset_filename>" by the name of your own dataset, e.g.:
  #   file = "dat/breslow_chatterjee_1999.csv",
  file = "dat/<dataset_filename>",
  col_types      = cols(
    instit       = col_factor(),
    histol       = col_factor(),
    stage        = col_factor(levels = as.character(1:4), ordered = TRUE),
    study        = col_factor(),
    in.subcohort = col_logical(),
    .default     = col_integer()
  )
)

## Recode values:
study_data <- study_data |> mutate(
  instit = instit |>
    factor(levels = 1:2, labels = c("Favourable", "Unfavourable")),
  histol = histol |>
    factor(levels = 1:2, labels = c("Favourable", "Unfavourable")),
  stage  = stage  |>
    factor(levels = 1:4, labels = c('I', 'II', 'III', 'IV'))
)

## Assign labels:
# (see https://www.rdocumentation.org/packages/survival/versions/3.5-5/topics/nwtco
#   for more info):
study_data <- study_data |> set_variable_labels(
  instit = "Histology (from local institution)",
  histol = "Histology (from central lab)",
  stage  = "Disease stage",
  study  = "Study",
  rel    = "Relapse", # As a 0/1 integer for compatibility with {survival}
  edrel  = "Time to relapse (days)",
  age    = "Age (months)"
  ## TODO: Transform `edrel` and `age` to years?
)


## Filter only cases in the example in Breslow & Chatterjee (1999):
study_data <- study_data |>
  filter(in.subcohort)   |>
  select(-in.subcohort)
```

```{r model-fit}
survival_fit <- coxph(Surv(edrel, rel) ~ histol + instit, data = study_data)
## TODO: Add covariates?
```

# Results

In the following we see the descriptive statistics and the contingency table.

```{r descriptives-table}
## TODO: Uncomment next line if using {gtsummary}
# study_data |> tbl_summary(include = -seqno)
```

```{r contingency-table}
## TODO: Uncomment next line if using {gtsummary}
# study_data |> tbl_cross(row = instit, col = histol)
```

The next table shows the coefficients of the model.

```{r model-output}
## TODO: Uncomment next line if using {gtsummary}
# survival_fit |> tbl_regression()
```
